---
title: "steamAPI"
output: html_document
date: "2022-12-23"
---

# R notebook to download games (apps) from Steam using their APIs

```{r}
library(tidyverse)
library(httr)
library(rvest)
library(lubridate)
```

In order to do the analysis, I'll first pull all the app IDs currently available on steam. As of. 12/23/2022 it is 152217. There are 54 apps that have a placeholder id (but they're not available to see for the public yet, so no name). I decided to eliminate those, leaving me with 152163 apps. There were also some other mistakes present, so the final number of apps is 152194.

## Initialize Tibble and populate with appIDs and names

> NOTE: The code below only needs to be run once; I commented it out once done.

```{r}
  # connect to the steam API with list of appIDs
    complete_app_list <- GET('http://api.steampowered.com/ISteamApps/GetAppList/v0002/',
      query = list(format='json'))

  # create tibble with info that will be collected
    steam_apps_df <- tibble(ID = numeric(),
                      name_tot = character(),
                      type = character(),
                      name_app = character(),
                      required_age = numeric(),
                      publishers = character(),
                      developers = character(),
                      is_free = logical(),
                      dlc = character(),
                      windows_mac_linux = character(),
                      metacritic = numeric(),
                      categories = character(),
                      genres = character(),
                      n_screenshots = numeric(),
                      n_trailers = numeric(),
                      recommended = numeric(),
                      achievements = numeric(),
                      release_date = date()
    )

    # this variable only used for printout to see how the construction of the tibble is going
    counting <- 0

    for (i in content(complete_app_list)$applist$apps) {
      if (i$name != '') {
        steam_apps_df <- steam_apps_df %>% 
          add_row(ID = i$appid, name_tot = i$name)
      }
      counting = counting + 1
      if (counting %% 15000 == 0){
        print(counting)
      }
    }


    saveRDS(steam_apps_df, file = "../data/complete_apps_list.rds")
```

## Get individual apps data

> NOTE: This next line is NOT needed if proceeding straight from above. But if any mistakes are made, you can always read in the last version instead of re-running the stuff.

```{r}
steam_apps_df <- readRDS('../data/complete_apps_list.rds')
```

The `get_app_details` function below is where the magic happens. I put comments inline, but essentially, the function takes the list of appIDs that don't have the column "name_app" filled (that column represents names taken directly from individual apps - sometimes formatted differently than "name_tot" obtained from the complete list).

Afterwards, it starts a for loop on those, calling the Steam API "appdetails" to fill out one row at a time - there's also a few cases where a row gets discarded.

```{r}
get_app_details <- function(apps_df, num_apps){
  list_ID <- steam_apps_df %>% filter(is.na(name_app)) %>% pull(ID)
  
  # loop over the list_ID of apps, excluding ones we have already as above
  # (equivalent of enumerate in python)
  for (i in seq_along(list_ID)) {
    
    # break when we reached the number of apps we want the details for
    if (i > num_apps) {
      break
    }
    
    # wait for 1.3 sec - steam API has limited # of requests
    # this seemed to be the fastest I could go without getting a 429 (too many requests) too often
    Sys.sleep(1.3)
    
    # at first used as.character(ID), but then a wild 6e05 messed it up,
    # so had to format and that transforms it to character!
    list_char_ID <- format(list_ID[[i]], scientific=FALSE)
    
    app_details <- GET('https://store.steampowered.com/api/appdetails/',
                      query = list(appids = list_char_ID, format='json'))
    
   
    if (status_code(app_details) != 200) {
      # break the loop if return code is not 200
      print(paste('uh oh ..', status_code(app_details), 'for id', list_ID[[i]]))
      break
    } else if (is.null(content(app_details))) {
      # found this after several thousand successes, sometimes app returns a null!
        apps_df <- apps_df %>%
        filter(ID != list_ID[[i]])
    } else if (content(app_details)[[list_char_ID]]$success == FALSE) {
      # if the app returns status false it means it's not there!
      apps_df <- apps_df %>%
        filter(ID != list_ID[[i]])
    } else {

      # to make the rest of this code easier to read, save to a tmp variable
      tmp_data <- content(app_details)[[list_char_ID]]$data
      
      # the following variables are obtained through loops since they're in a list of variable size
      tmp_developers <- ''
      tmp_publishers <- ''
      tmp_dlc <- ''
      tmp_categories <- ''
      tmp_genres <- ''
      
      for (developer in tmp_data$developers) {
        tmp_developers <- paste(tmp_developers, developer, sep=';')}
      for (publisher in tmp_data$publishers) {
        tmp_publishers <- paste(tmp_publishers, publisher, sep=';')}
      for (dlc in tmp_data$dlc) {
        tmp_dlc <- paste(tmp_dlc, dlc, sep=';')}
      for (category in tmp_data$categories) {
        tmp_categories <- paste(tmp_categories, category$description, sep=';')}
      for (gen in tmp_data$genres) {
        tmp_genres <- paste(tmp_genres, gen$description, sep=';')}
      
      # finally, getting all of the data in the tibble
      apps_df <- apps_df %>% 
        rows_update(tibble(
                ID = list_ID[[i]],
                type = tmp_data$type,
                name_app = tmp_data$name,
                required_age = as.numeric(tmp_data$required_age),
                publishers = na_if(str_remove(tmp_publishers, ';'), ''),
                developers = na_if(str_remove(tmp_developers, ';'), ''),
                is_free = tmp_data$is_free,
                dlc = na_if(str_remove(tmp_dlc, ';'), ''),
                windows_mac_linux = paste(tmp_data$platforms$windows, tmp_data$platforms$mac, tmp_data$platforms$linux, sep=';'),
                metacritic = tmp_data$metacritic$score,
                categories = na_if(str_remove(tmp_categories, ';'), ''),
                genres = na_if(str_remove(tmp_genres, ';'), ''),
                n_screenshots = length(tmp_data$screenshots),
                n_trailers = length(tmp_data$movies),
                recommended = tmp_data$recommendations$total,
                achievements = tmp_data$achievements$total,
                release_date = tmp_data$release_date$date),
              by = 'ID')
      
    }  
  }
  
  # after the wanted number of iterations, return the updated tibble
  return(apps_df)
}

```

With the function defined, all it remains is to run through all of the apps individually. I was doing it in chunks to perform some additional checks as I went through it (and there were some minor fixes I had to adjust). I chose the function to run through 100 games and then repeat with the desired number of loops, printing out occasionally.

> NOTE: That said, the below could just be ran through the entire list, however, it will take approx. 600 hours to do so and there ARE going to be some 429 responses every once in a while.

```{r}
for (i in 1:100) {
  if (i %% 10 == 0) {
    print(paste('at time:', now(), 'on outer loop number:',i))
  }
  steam_apps_df <- get_app_details(steam_apps_df, 100)
}
```

## Additional checks and saving

This next line of code simply used to check the progress (i.e. how many apps do I have the details on so far).

```{r}
count(steam_apps_df %>% filter(!is.na(name_app)))
```

The code below is to check for a particular slice (if something is going on)

```{r}
steam_apps_df %>% 
  slice(16800:17000) 
```

Finally, and perhaps most importantly, save the data so that it can be

```{r}
saveRDS(steam_apps_df, file = "../data/complete_apps_list_138746.rds")
```

## Ge publishers and developers from wikipedia

I decided to add this part to the steamAPI document, since it contains more webscraping. Obtaining the list of publishers and developers from Wikipedia.

```{r}
publisher_list <-  read_html('https://en.wikipedia.org/wiki/List_of_video_game_publishers')

publisher_list %>% 
  html_nodes('table') %>% 
  pluck(2) %>% 
  html_table %>% 
  saveRDS(file='../data/publisher_list.rds')


developer_list <-  read_html('https://en.wikipedia.org/wiki/List_of_video_game_developers')

developer_list %>% 
  html_nodes('table') %>% 
  pluck(4) %>% 
  html_table %>% 
  saveRDS(file='../data/developer_list.rds')

```

