---
title: "TU3_advanced_graphs"
output: html_document
date: "2023-01-11"
---

# More complex graphs prep for the shiny app

```{r}
library(tidyverse)
library(plotly)
library(igraph)
```

## reading in the data and filtering to get just the games

> NOTE: I might expand on this at a later date to include dlcs or other categories

```{r}
steam_apps_df <- readRDS('../data/complete_apps_list_FINAL.rds') %>% 
  distinct(name_app, .keep_all= TRUE)
```

```{r}
steam_games_df <- steam_apps_df%>% 
  filter(type=='game') %>% 
  select(-c('name_tot','type', 'required_age'))
```

Added later when I finalized the columns that I'm going to use

```{r}
steam_games_df <- steam_games_df %>% 
  mutate(release_year=str_extract(release_date, '\\d{4}')) %>% 
  mutate(release_year=replace(release_year, is.na(release_year) |
                                release_year < 1997 |
                                release_year > 2023, '20XY'))
```


## Network Graph - Categories

Checking out possible connections between categories for a network graph.

> I commented the code below because it takes a few min to run and ultimately I decided that there
was not much value in looking at the categories and their inter connectivity

```{r}
# Get the list of distinct categories
#categories_list <- steam_games_df %>% 
#  filter(!is.na(categories)) %>%
#  separate_rows(categories, sep = ';') %>% 
#  distinct(categories) %>% 
#  pull()

# dummy variable to count all different connections between pairs of categories
#count_connections <- 0

# loop over categories
# NOTE: this takes a few minutes to run
#for (j in 1:(length(categories_list)-1)) {
#  for (i in (j+1):length(categories_list)) {
#    if(steam_games_df %>% 
#       filter(str_detect(categories, categories_list[j]) &
#              str_detect(categories, categories_list[i])) %>% 
#       count() %>% 
#       pull()
#       != 0 ) {count_connections = count_connections + 1} 
#  }
#}
#
#print(count_connections)
```

I'd need 659 connections to map out categories in a network graph

## Network Graph - Genres

Checking out possible combinations for potentially building a network graph. This turned out to be a much better idea than categories.

```{r}
# counting apps by genre (overall)
apps_count_by_genre <- steam_games_df %>% 
  filter(!is.na(genres)) %>%
  separate_rows(genres, sep = ';') %>% 
  select(name_app, genres) %>%
  count(genres, sort=TRUE, name='num_apps') 

# same as above but having a breakdown by years
apps_count_by_genre_year <- steam_games_df %>% 
  filter(!is.na(genres)) %>%
  separate_rows(genres, sep = ';') %>% 
  select(name_app, release_year, genres) %>%
  count(genres, release_year, sort=TRUE, name='num_apps')

# how many years there are ..
years_list <- steam_games_df %>% 
  distinct(release_year) %>% 
  pull()
```

> NOTE: code commented below is the original version without breakdown by years.

```{r}
#different_connections <- 0
#count_connections <- 0

#relations <- tibble(from=character(),
#                    to=character(),
#                    num_connections=numeric()
#)

#for (j in 1:(nrow(apps_count_by_genre)-1)) {
#  for (i in (j+1):nrow(apps_count_by_genre)) {
#    count_connections <- steam_games_df %>% 
#      filter(str_detect(genres, (apps_count_by_genre %>% pull(genres))[j]) &
#               str_detect(genres, (apps_count_by_genre %>% pull(genres))[i])) %>% 
#      count() %>% 
#      pull()

#    if(count_connections != 0 ) {
#      different_connections = different_connections + 1
#      relations <- relations %>% 
#        add_row(from=(apps_count_by_genre %>% pull(genres))[j],
#                to= (apps_count_by_genre %>% pull(genres))[i],
#                num_connections=count_connections)
#    } 
#  }
#}

#print(different_connections)


```

If I include all genres, I'll need 389 connections. However. Cat sudoku is the only one that has like 3 categories that no other game has. I also started doing some testing. If I filter out any genres with less than 10 connections between, I loose 196 connections! 224 filtering on less than 20, 242 at less than 30, 261 at less than 40, 273 less than 50. 301 connections at less than 100.

Also with less than 10 counts of connections, I loose 6 genre categories right away., 20 loose 6; then have to go up to 80 to loose 7, 90 to loose 11. So filtering on 100 (just to have a round number instead of 90), will alow me to get rid of 11 genres that are very rarely represented.

Looking below, by cutting at less than 31 connections, I only loose 0.495% of overall number of connections (suggested possible cutoff at 0.5% by Michael). With it, I lose the following 7 genres: Accounting (24 games), Movie (2 games), 360 Video (1 game), Documentary (1 game), Episodic (1 game), Short (1 game), Tutorial (1 game).

I also did some additional testing (successively deleted the code since i decided against it) where I either cut our certain "weak" connections as a small percentage of the total, or, weighted them (depending on the possible size - in the shiny app I included a table with this later info).

Ultimately, I decided that I would not make any cuts and instead, do a breakdown over release years.

> NOTE: the code below takes about 20 minutes to run!!

```{r}
different_connections <- 0
count_connections <- 0

relations <- tibble(from=character(),
                    to=character(),
                    num_connections=numeric(),
                    release_year=character()
)


for (yr in years_list) {
  games_by_year <- steam_games_df %>% 
    filter(release_year == yr)
  
  for (j in 1:(nrow(apps_count_by_genre)-1)) {
    for (i in (j+1):nrow(apps_count_by_genre)) {
      count_connections <- games_by_year %>% 
        filter(str_detect(genres, (apps_count_by_genre %>% pull(genres))[j]) &
                 str_detect(genres, (apps_count_by_genre %>% pull(genres))[i])) %>% 
        count() %>% 
        pull()
      
      if(count_connections != 0 ) {
        different_connections = different_connections + 1
        relations <- relations %>% 
          add_row(from=(apps_count_by_genre %>% pull(genres))[j],
                  to= (apps_count_by_genre %>% pull(genres))[i],
                  num_connections=count_connections,
                  release_year=yr)
      } 
    }
  }
}

print(different_connections)

# below to save
#relations %>%  saveRDS('../shiny_app/data/genres_relations_by_year.rds')

```

## Work in progress - a more efficient way to count connections

After speaking with Michael, trying to do a more efficient count rather than a triple for loop..

The idea being:
1) separating genres into rows as before
2) filling a 'dummy' column with 1
3) pivoting table to individual genres columns (up to 33 of them) and filling missing values with 0
4) counting that matrix

> Tested something fairly obvious: if I separate rows in genres first, and filter later, the proces
is slower than if I filter first and then separate later. However, if I keep reiterating (aka changing
the filter conditions), then the approach of separating all rows for all years is better.

```{r}
# Start the clock!
ptm <- proc.time()

steam_games_df %>% 
  filter(!is.na(genres)) %>%
  separate_rows(genres, sep = ';') %>% 
  select(name_app, release_year, genres) %>% 
  add_column(dummy=1) %>% 
  pivot_wider(id_cols=c('name_app', 'release_year'),
              names_from='genres', values_from='dummy',
              values_fill=0)


# Stop the clock
proc.time() - ptm

```

```{r}
distinct_genres <- steam_games_df %>% 
  filter(!is.na(genres)) %>%
  separate_rows(genres, sep = ';') %>% 
  distinct(genres) %>%
  pull()

expand_grid(x=distinct_genres, y=distinct_genres)
permutations(n=33,r=2,v=distinct_genres, repeats.allowed=FALSE)
```


## Drawing the Genres Network Graph

```{r}
# to make the igraph
#g <- graph_from_data_frame(relations_filtered, directed=FALSE, vertices=apps_count_by_genre_filtered)
vertices_filtered <- apps_count_by_genre_year_filtered %>% 
  filter(release_year==2006) %>% 
  group_by(genres) %>% 
  summarize(num_apps=sum(num_apps)) %>% 
  arrange(desc(num_apps))

relations_filtered <- relations_cut %>% 
  filter(release_year==2006) %>% 
  group_by(from, to) %>% 
  summarize(num_connections=sum(num_connections)) %>% 
  ungroup()

g <- graph_from_data_frame(relations_filtered, directed=FALSE, vertices=vertices_filtered )
```

```{r}
G <- upgrade_graph(g)
#L <- layout_in_circle(G)
#set.seed(300)
L <- layout_nicely(G)
#L <- layout_with_dh(G)

#L <- candidate_layout_01

vs <- V(G)
es <- as.data.frame(get.edgelist(G))
Nv <- length(vs)
Ne <- length(es[1]$V1)

Xn <- L[,1]
Yn <- L[,2]

#unique_names <- unique(c(es$V1, es$V2))

unique_names <- unique(vertices_filtered$genres)
names(Xn) <- unique_names
names(Yn) <- unique_names


network <- plot_ly(x = ~Xn, 
                   y = ~Yn, 
                   mode = 'markers', 
                   marker = list(size = 100* (vertices_filtered %>% pull(num_apps)) / (vertices_filtered %>% select(num_apps) %>% sum()),
                                 opacity = 0.7),
                   #text = vs$label, 
                   text = paste(unique_names, '<br>', vertices_filtered$num_apps),
                   texttemplate = if_else(100* (vertices_filtered %>% pull(num_apps)) / (vertices_filtered %>% select(num_apps) %>% sum()) >10, paste(unique_names, '<br>', vertices_filtered$num_apps), ''),
                   hoverinfo = 'text'
) %>% 
  add_text(textposition= 'right')


edge_shapes <- list()

for(i in 1:Ne) {
  v0 <- es[i,]$V1
  v1 <- es[i,]$V2
  edge_shape = list(
    type = 'line',
    line = list(color = "#030303",
                width = 30 * (relations_filtered %>% pull(num_connections))[i] / (relations_filtered %>% select('num_connections') %>% sum())
    ),
    x0 = Xn[v0],
    y0 = Yn[v0],
    x1 = Xn[v1],
    y1 = Yn[v1]
  )
  edge_shapes[[i]] <- edge_shape
}

axis <- list(title = "", showgrid = FALSE, showticklabels = FALSE, zeroline = FALSE)


fig <- layout(
  network,
  title = 'Genres Network',
  shapes = edge_shapes,
  xaxis = axis,
  yaxis = axis
)

fig

```


## Some Tests for Shiny App ...

Below the two different variants. I first started with plotly as I thought the interactivity would be useful. But then, with the filters that I was able to add to the shiny app, I changed from plotly to ggplot2 for efficiency.

```{r}
steam_games_df %>%
  mutate(release_type=if_else(release_year == 'unknown' | release_year == 2023, 'Unreleased', 'Released')) %>%
  count(release_year, release_type, name='num_games') %>% 
  plot_ly(
    x = ~release_year,
    y = ~num_games,
    text = ~num_games,
    textposition = 'outside',
    texttemplate = ~if_else(num_games >= 1000, '%{y:.2s}', '%{y:.1s}'),
    hovertemplate = ~paste('Year:', release_year, '<br>N. Games:', num_games),
    name = ~release_type,
    type = 'bar',
    color = ~release_type,
    colors = c('#367fa9', '#1b2838')
  ) %>% 
  layout(title = 'Video Game Releases by Year',
         xaxis = list(title = 'Release Year'),
         yaxis = list(title = 'Number of Video Games'),
         legend = list(x = 0.1, y = 0.9)
  )
```
```{r}
steam_games_df %>%
  mutate(release_type=if_else(release_year == '20XY' | release_year == 2023, 'Unreleased', 'Released')) %>%
  #filter(release_type != 'Unreleased') %>% 
  #filter(release_year < 2013) %>% 
  count(release_year, release_type, name='num_games') %>% 
  ggplot(aes(x=release_year, y=num_games, fill=release_type, label=format(num_games, big.mark=',') )) +
  ggtitle('Video Game Releases by Year') +
  scale_fill_manual(values=c('#367fa9', '#1b2838')) +
  scale_color_manual(values=c('#367fa9', '#1b2838')) +
  geom_col() +
  geom_text(size=2.5, vjust=-0.5, aes(color=release_type)) +
  labs(x ='Release Year', y = 'Number of Video Games', fill='Game Status', color='Game Status') +
  #geom_vline(xintercept = '2014', linetype="dashed") +
  #  annotate("text", x='2014', y= 5000, label='World Average', angle = 90) +
  theme(axis.text.x=element_text(angle = 90, hjust = 0, vjust=0), legend.position=c(0.15, 0.8))

```

```{r}
# Start the clock!
ptm <- proc.time()

steam_games_df %>% 
  filter(!is.na(genres)) %>%
  separate_rows(genres, sep = ';') %>% 
  select(name_app, release_year, genres) %>% 
  saveRDS('test.rds')

# Stop the clock
proc.time() - ptm
```


```{r}
genres_relations_df %>% 
  group_by(from, to) %>% 
  summarize(total_connections=sum(num_connections)) %>% 
  ungroup() %>% 
  arrange(desc(total_connections)) %>% 
  left_join(steam_games_df %>% 
              separate_rows(genres, sep = ';') %>% 
              drop_na(genres) %>% 
              count(genres, sort=TRUE, name='num_apps'),
            by=c('from'='genres'), suffix=c('_from', '_to')) %>% 
  left_join(steam_games_df %>% 
              separate_rows(genres, sep = ';') %>% 
              drop_na(genres) %>% 
              count(genres, sort=TRUE, name='num_apps'),
            by=c('to'='genres'), suffix=c('_from', '_to')) %>% 
  mutate(possible_connections= paste0(format(100*total_connections / pmin(num_apps_from, num_apps_to), digits=1), '%')) %>% 
  select(from, to, num_apps_from, num_apps_to, total_connections, possible_connections)


steam_games_df %>% 
  mutate(webpage= paste0('https://store.steampowered.com/app/', ID))
```

## Shiny App Optimization

After presenting my app on 1/21/2023, I wanted to optimize the code. Namely, I wanted the app to load faster (both on my laptop, as well as on shinyapps.io). So I did a bunch of testing with elapsed time and ultimately broke down the dataframe into multiple so that the code would execute quickly.

I was able to significantly improve load times, making it especially noticeable when changing the filtering options within the app - the refresh time went from a couple of seconds to a fraction of one!

```{r}
# games dataset with columns that don't need any additional string manipulations
# changed columns from char to logical where possible (since I ended using only that info)
steam_games_df %>%
  separate(windows_mac_linux, c('is_win', 'is_mac', 'is_linux'), sep = ';') %>%
  mutate(is_win = as.logical(is_win),
         is_mac = as.logical(is_mac),
         is_linux = as.logical(is_linux),
         has_dlc=ifelse(is.na(dlc), FALSE, TRUE),
         has_recommended=ifelse(is.na(recommended), FALSE, TRUE),
         has_metacritic=ifelse(is.na(metacritic), FALSE, TRUE),
         has_achievements=ifelse(is.na(achievements), FALSE, TRUE)) %>%
  select('ID', 'name_app', 'is_free', 'is_win', 'is_mac', 'is_linux', 'has_dlc', 'has_recommended', 'has_metacritic', 'has_achievements', 'release_date', 'release_year', 'n_screenshots', 'n_trailers') %>% 
  saveRDS('../shiny_app/data/filtered_games.rds')

steam_games_df %>% 
  drop_na(publishers) %>% 
  separate_rows(publishers, sep = ';') %>% 
  select(ID, publishers, release_year) %>% 
  saveRDS('../shiny_app/data/publishers_by_year.rds')

steam_games_df %>% 
  drop_na(developers) %>% 
  separate_rows(developers, sep = ';') %>% 
  select(ID, developers, release_year) %>% 
  saveRDS('../shiny_app/data/developers_by_year.rds')

steam_games_df %>% 
  drop_na(genres) %>% 
  separate_rows(genres, sep = ';') %>% 
  group_by(release_year) %>% 
  count(genres, sort=TRUE, name='num_apps') %>% 
  ungroup() %>%
  saveRDS('../shiny_app/data/genres_by_year.rds')

steam_games_df %>% 
  drop_na(categories) %>% 
  separate_rows(categories, sep = ';') %>% 
  group_by(release_year) %>% 
  count(categories, sort=TRUE, name='num_apps') %>% 
  ungroup() %>% 
  saveRDS('../shiny_app/data/categories_by_year.rds')

```




1. fork code
2. delete old one
3. pull down new one

1. new repo personal github
2. git remote add
