---
title: "TU3_advanced_graphs"
output: html_document
date: "2023-01-11"
---

# More complex graphs prep for the shiny app

```{r}
library(tidyverse)
library(plotly)
library(igraph)
```

## reading in the data and filtering to get just the games

> note: I might expand on this at a later date

```{r}
steam_apps_df <- readRDS('../data/complete_apps_list_FINAL.rds') %>% 
  distinct(name_app, .keep_all= TRUE)
```

```{r}
steam_games_df <- steam_apps_df%>% 
  filter(type=='game') %>% 
  select(-type)
```

## Network Graph - Categories

Checking out possible connections between categories for a network graph.

```{r}
categories_list <- steam_games_df %>% 
    filter(!is.na(categories)) %>%
    separate_rows(categories, sep = ';') %>% 
    distinct(categories) %>% 
    pull()
```

```{r}
count_connections <- 0

for (j in 1:(length(categories_list)-1)) {
  for (i in (j+1):length(categories_list)) {
    if(steam_games_df %>% 
       filter(str_detect(categories, categories_list[j]) &
              str_detect(categories, categories_list[i])) %>% 
       count() %>% 
       pull()
       != 0 ) {count_connections = count_connections + 1} 
  }
}

print(count_connections)
```

I'd need 659 connections to map out categories in a network graph

## Network Graph - Genres

Checking out possible combinations for potentially building a network graph. This turned out to be a much better idea than categories.

```{r}
apps_count_by_genre <- steam_games_df %>% 
    filter(!is.na(genres)) %>%
    separate_rows(genres, sep = ';') %>% 
    select(name_app, genres) %>%
    count(genres, sort=TRUE, name='num_apps') 
```

```{r}
different_connections <- 0
count_connections <- 0

relations <- tibble(from=character(),
                    to=character(),
                    num_connections=numeric()
)

for (j in 1:(nrow(apps_count_by_genre)-1)) {
  for (i in (j+1):nrow(apps_count_by_genre)) {
    count_connections <- steam_games_df %>% 
       filter(str_detect(genres, (apps_count_by_genre %>% pull(genres))[j]) &
              str_detect(genres, (apps_count_by_genre %>% pull(genres))[i])) %>% 
       count() %>% 
       pull()
    
    if(count_connections != 0 ) {
      different_connections = different_connections + 1
      relations <- relations %>% 
        add_row(from=(apps_count_by_genre %>% pull(genres))[j],
                to= (apps_count_by_genre %>% pull(genres))[i],
                num_connections=count_connections)
    } 
  }
}

print(different_connections)


```

If I include all genres, I'll need 389 connections. However. Cat sudoku is the only one that has like 3 categories that no other game has. I also started doing some testing. If I filter out any genres with less than 10 connections between, I loose 196 connections! 224 filtering on less than 20, 242 at less than 30, 261 at less than 40, 273 less than 50. 301 connections at less than 100.

Also with less than 10 counts of connections, I loose 6 genre categories right away., 20 loose 6; then have to go up to 80 to loose 7, 90 to loose 11. So filtering on 100 (just to have a round number instead of 90), will alow me to get rid of 11 genres that are very rarely represented.

Looking below, by cutting at less than 31 connections, I only loose 0.495% of overall number of connections (suggested possible cutoff at 0.5% by Michael). With it, I lose the following 7 genres: Accounting (24 games), Movie (2 games), 360 Video (1 game), Documentary (1 game), Episodic (1 game), Short (1 game), Tutorial (1 game)

```{r}
relations %>% 
  filter(num_connections < 32) %>% 
  summarize(sum(num_connections)) %>% 
  pull() / relations %>% summarize(sum(num_connections)) %>% pull()

```


```{r}
rel_from <- relations %>% 
  filter(num_connections >= 32) %>% 
  distinct(from)

rel_to <- relations %>% 
  filter(num_connections >= 32) %>% 
  distinct(to)

genres_filtered <- rel_from %>% 
  full_join(rel_to, by=c('from'='to')) %>% 
  pull()

```

```{r}
apps_count_by_genre_filtered <- apps_count_by_genre %>% 
  filter(genres %in% genres_filtered)

relations_filtered <- relations %>% 
  filter(num_connections >= 32)
```

```{r}
# attempted to weight the connections, didn't yield results I wanted
#relations_weighted <- relations_filtered %>%
#  left_join(apps_count_by_genre_filtered, by=c('from'='genres')) %>% 
#  rename(from_num_apps=num_apps) %>% 
#  left_join(apps_count_by_genre_filtered, by=c('to'='genres')) %>% 
#  rename(to_num_apps=num_apps) %>% 
#  mutate(adjusted_connections= num_connections / pmin(from_num_apps, to_num_apps)) %>% 
#  select(from, to, adjusted_connections)

# also have to add this code below
# width = if_else((relations_weighted %>% pull(adjusted_connections))[i] > 0.5, (relations_weighted %>% pull(adjusted_connections))[i], 0)
```


Time to do the network graph!!

```{r}
# to make the igraph
g <- graph_from_data_frame(relations_filtered, directed=FALSE, vertices=apps_count_by_genre_filtered)
```

```{r}
G <- upgrade_graph(g)
#L <- layout_in_circle(G)
#set.seed(300)
#L <- layout_nicely(G)
#L <- layout_with_dh(G)

L <- candidate_layout_01

vs <- V(G)
es <- as.data.frame(get.edgelist(G))
Nv <- length(vs)
Ne <- length(es[1]$V1)

Xn <- L[,1]
Yn <- L[,2]

unique_names <- unique(c(es$V1, es$V2))
names(Xn) <- unique_names
names(Yn) <- unique_names

network <- plot_ly(x = ~Xn, 
                   y = ~Yn, 
                   mode = 'markers', 
                   marker = list(size = 100* (apps_count_by_genre_filtered %>% pull(num_apps)) / (apps_count_by_genre_filtered %>% select(num_apps) %>% sum()),
                                 opacity = 0.7),
                   #text = vs$label, 
                   text = paste(unique_names, '<br>', apps_count_by_genre_filtered$num_apps),
                   texttemplate = if_else(100* (apps_count_by_genre_filtered %>% pull(num_apps)) / (apps_count_by_genre_filtered %>% select(num_apps) %>% sum()) >10, paste(unique_names, '<br>', apps_count_by_genre_filtered$num_apps), ''),
                   hoverinfo = 'text'
                   ) %>% 
  add_text(textposition= 'right')


edge_shapes <- list()

for(i in 1:Ne) {
  v0 <- es[i,]$V1
  v1 <- es[i,]$V2
  edge_shape = list(
    type = 'line',
    line = list(color = "#030303",
                width = 30 * (relations_filtered %>% pull(num_connections))[i] / (relations_filtered %>% select('num_connections') %>% sum())
                ),
    x0 = Xn[v0],
    y0 = Yn[v0],
    x1 = Xn[v1],
    y1 = Yn[v1]
  )
  edge_shapes[[i]] <- edge_shape
}

axis <- list(title = "", showgrid = FALSE, showticklabels = FALSE, zeroline = FALSE)


fig <- layout(
  network,
  title = 'Genres Network',
  shapes = edge_shapes,
  xaxis = axis,
  yaxis = axis
)

fig



```